Demystifying the Combination of Dynamic Slicing and \\ Spectrum-based Fault Localization

### 2019-02-22 paper/ijcai/paper.tex
organizing and polishing text (moving to eval)

Given the importance of fault localization, this paper revisits the
problem of assessing the effectiveness of combining \ds{} and \sfl{},
addressing the issues from prior work. The study we conducted involves
\numFaults{} faults from \numPrograms{}
different programs from the Defects4J (D4J)
dataset~\cite{just-defects4j-issta2014}, which is frequently used to
evaluate fault localization research. The study covers the following
research questions:

\newcommand{\rqone}{How often does DS miss faulty statements?}
\newcommand{\rqthree}{How effective is the \comb{} approach?}

\begin{itemize}[leftmargin=*]
\item[]{\footnotesize[(un)soundness]}~\textit{\rqone{}}
\item[]{\footnotesize[effectiveness]}~\textit{\rqthree{}}
\end{itemize}

The first question addresses the problem that affects all dynamic
slicing techniques~\cite{Lin:2018:BDE:3238147.3238163}--that of
missing faulty statements. Naturally, the consequence of that problem
for debugging is that developers would never be able to locate
faults. Different slicers can miss faulty statements for different
reasons. In this paper, we used Critical
Slicing~\cite{DeMillo:1996:CSS:229000.226310}, which, in principle,
could miss statements because of imprecise oracles that guide the
slicing process.\comments{ We evaluated how often that issue manifests
  in our experiments.} The second research question evaluates the
impact of the combination. To sum, our results show that the
combination of the two approaches improves the accuracy of fault
localization on average by \avgImprov\ in 51\% of the studied faults.

The contributions of this work are: 1)~an empirical study, using
real-world applications and bugs, on the combination of \ds{} and
\sfl{} for bug localization of \emph{Java} faulty programs and 2)~a
tool that implements the combination of \ds{} with \sfl{}, coined
\comb{}.  The tool as well as a replication package will be available
once double-blind requirements are lifted.

@@

\section{Background}
\label{sec:background}

This section summarizes the ground concepts of our approach.

\subsection{Dynamic Slicing (\ds{})}
\label{sec:slicing}

@@

\subsection{Spectrum-based Fault Localization (\sfl)}
\label{sec:sfl}

@@

Spectrum-based fault localization is a statistical fault
localization technique that takes as input a test suite including at
least one failing test and reports on output a ranked list of
components likely to be in
fault~\cite{FLSurvey2016,DBLP:conf/kbse/JonesH05,DBLP:journals/smr/LuciaLJTB14,DBLP:journals/jss/AbreuZGG09}. The
following are given in \sfl{}: a finite set $\mathcal{C}=\set{c_1,c_2,...,c_M}$
of $M$ system \emph{components}\footnote{A
component can be any code artifact of arbitrary granularity
such as a class, a method, or a statement~\cite{DBLP:journals/stvr/HarroldRSWY00}.};
a finite set $\mathcal{T} = \set{t_1,t_2,...,t_N}$ of $N$ system transactions,
which correspond to records of a system execution, such as test cases;
the error vector $e = \set{e_1,e_2,...,e_N}$, where $e_i = 1$ if
transaction $t_i$ has failed and $e_i = 0$ otherwise; and an
$N\times{}M$ coverage matrix $\mathcal{A}$, where $\A{ij}$ denotes the
coverage of component $c_j$ in transaction $t_i$.  The pair
$(\mathcal{A},e)$ is commonly referred to as
spectrum~\cite{DBLP:journals/stvr/HarroldRSWY00}. Figure~\ref{fig:spectrum-example}
shows an example spectrum.

@@

\section{The \comb{} Approach}
\label{sec:approach}

This section provides details on the proposed approach, \combpar{}, that combines
\ds{} with \sfl{}; it also discusses the oracle problem.

\subsection{Combining \ds{} and \sfl{}}
\label{sec:comb}

To illustrate the idea behind the combination of \sfl{} with \ds{}, consider a
debugging scenario with five components $c_{1..5}$ and five
transactions $t_{1..5}$, two of which are failing. Consider,
additionally, that the component $c_2$ is faulty.
Figure~\ref{fig:illustration} shows, at the left-hand side of the
arrow ($\Rightarrow$), an hypothetical spectra and its corresponding
ranking, produced with the Ochiai predictor. Each line in the ranking
shows, respectively, the rank/position, the component label, and the
Ochiai score (in parentheses).

@@

In the following, we illustrate the workflow of the \comb{}
approach for this scenario. \comb{} uses a granularity of statements.
First, the combined technique picks statements $c_4$ and $c_3$, which
appear at the top of the ranking, for further analysis.
(The cutoff point is user-defined.) Second, the
technique picks a failing test $t_2$ to slice the code. A slice is a
set of statements. Let us assume
that the slice obtained for the transaction $t_2$ is $\{c_2, c_5\}$,
\ie, it excludes statements $c_4$ and $c_3$. Finally, the spectra and
ranking are updated. The right-side of Figure~\ref{fig:illustration}
shows the modified spectra and ranking after slicing test $t_2$
against $c_4$ and $c_3$. It is worth noting that the technique slices
the code for every failing tests. More precisely, the workflow of the
\comb{} approach consists of four steps:
\begin{enumerate}
  \itemsep0em
  \item\label{step:spectra} Compute spectra $S$ and ranking $R$ for
  the input test suite;
  \item\label{step:susp-files} Select top $k$ most suspicious classes, according to $R$;
  \item\label{step:compute-slice} Compute slicer for every suspicious
    file, obtained in Step~\ref{step:susp-files}, and every failing test;
  \item\label{step:group} Assemble all the resulting slicers of each failing test;
  \item Adjust the spectra $S$, obtained in Step~\ref{step:spectra}, with
    the slices of each failing test obtained in Step~\ref{step:group} and
	recompute ranking.
\end{enumerate}

Intuitively, slicing enables the identification of statements in the spectra
whose values can be set to zero, \ie{}, the components marked with $0$
are irrelevant to determine the test output. Let us now observe the
effect of this spectra modification on the ranking. Analyzing the
Ochiai formula, one observes that, for the components $j$ which are
not part of the slice of a failing test, the combination reduces the
value of $\N{11}$ and increases the value of $\N{01}$. Therefore,
suspiciousness of those components decrease. Similar argument applies
to other fault predictors. In the running example,
Figure~\ref{fig:illustration}, components $c_3$ and $c_4$, which are
not in the slice of $t_2$, have their suspiciousness reduced, enabling
the faulty component $c_2$ to rise from the third to the first
position in the ranking.

@@

%
\section{Evaluation}
\label{sec:eval}
%
To evaluate our approach, we compare the cost of diagnosing
a collection of faults using \sfl{} versus \comb{}.
%
\subsection{Objects of Analysis}\label{sec:analysis}

>>>>

Given the importance of fault localization, this paper revisits the
problem of assessing the effectiveness of combining \ds{} and \sfl{},
addressing the issues from prior work. The study we conducted involves
\numFaults{} faults from \numPrograms{}
different programs from the Defects4J (D4J)
dataset~\cite{just-defects4j-issta2014}, which is frequently used to
evaluate fault localization research. The contributions of this work
are: 1)~an empirical study, using real-world applications and bugs, on
the combination of \ds{} and \sfl{} for bug localization of
\emph{Java} faulty programs and 2)~a tool that implements the
combination of \ds{} with \sfl{}, coined \comb{}.  The tool as well as
a replication package will be available once double-blind requirements
are lifted.

@@

%\section{Background}
%This section summarizes the ground concepts of our approach.

\section{Dynamic Slicing (\ds{})}
%\label{sec:background}
\label{sec:slicing}

@@

\section{Spectrum-based Fault Localization (\sfl)}
\label{sec:sfl}

@@

Spectrum-based fault localization is a statistical fault
localization technique that takes as input a test suite including at
least one failing test and reports on output a ranked list of
components likely to be in
fault~\cite{7390282,DBLP:conf/kbse/JonesH05,DBLP:journals/smr/LuciaLJTB14,DBLP:journals/jss/AbreuZGG09}. The
following are given in \sfl{}: a finite set $\mathcal{C}=\set{c_1,c_2,...,c_M}$
of $M$ system \emph{components}\footnote{A
component can be any code artifact of arbitrary granularity
such as a class, a method, or a statement~\cite{DBLP:journals/stvr/HarroldRSWY00}.};
a finite set $\mathcal{T} = \set{t_1,t_2,...,t_N}$ of $N$ system transactions,
which correspond to records of a system execution, such as test cases;
the error vector $e = \set{e_1,e_2,...,e_N}$, where $e_i = 1$ if
transaction $t_i$ has failed and $e_i = 0$ otherwise; and an
$N\times{}M$ coverage matrix $\mathcal{A}$, where $\A{ij}$ denotes the
coverage of component $c_j$ in transaction $t_i$.  The pair
$(\mathcal{A},e)$ is commonly referred to as
spectrum~\cite{DBLP:journals/stvr/HarroldRSWY00}. Figure~\ref{fig:spectrum-example}
shows an example spectrum.

@@

\section{The \comb{} Approach}
\label{sec:approach}
\label{sec:comb}

This section describes \combpar{}, a technique where \ds{} and \sfl{}
work in tandem to solve the fault localization problem. To illustrate
the idea behind \combpar{}, consider a debugging scenario with five
components $c_{1..5}$ and five transactions $t_{1..5}$, two of which
are failing. Consider, additionally, that the component $c_2$ is
faulty.  Figure~\ref{fig:illustration} shows, at the left-hand side of
the arrow ($\Rightarrow$), an hypothetical spectra and its
corresponding ranking, produced with the Ochiai predictor. Each line
in the ranking shows, respectively, the rank/position, the component
label, and the Ochiai score (in parentheses).

@@

We illustrate the workflow of the \comb{} approach for this scenario
in the following. Let us analyze components at the granularity of
statements.  First, \comb{} picks the first two statements at the top
of the ranking, $c_4$ and $c_3$, for further analysis.  (The cutoff
point is user-defined.) Second, the technique picks a failing test to
slice the code, say $t_2$. A slice is a set of statements. Let us
assume that the slice obtained for the transaction $t_2$ is $\{c_2,
c_5\}$, \ie, it excludes statements $c_4$ and $c_3$ that were
previously covered by the test. Finally, the spectra and ranking are
updated. The right-side of Figure~\ref{fig:illustration} shows the
modified spectra and ranking after slicing the code against $t_2$. It
is worth noting that the technique slices the code for every failing
tests. Intuitively, slicing enables the identification of statements
in the spectra whose values can be set to zero, \ie{}, the components
marked with $0$ are irrelevant to determine the test output.

Let us now observe the effect of this spectra modification on the
ranking. Analyzing the Ochiai formula, one observes that, for the
components $j$, which are not part of the slice of a failing test, the
combination reduces the value of $\N{11}$ and increases the value of
$\N{01}$. Therefore, suspiciousness of those components
decrease. Similar argument applies to other fault predictors. In the
running example from Figure~\ref{fig:illustration}, the components
$c_3$ and $c_4$, which are not in the slice of $t_2$, have their
suspiciousness reduced, enabling the faulty component $c_2$ to rise
from the third to the first position in the ranking.

\vspace{1ex}
To sum, the workflow of \comb{} consists of four steps:
\begin{enumerate}
  \itemsep0em
  \item\label{step:spectra} Compute spectra $S$ and ranking $R$ for
  the input test suite;
  \item\label{step:susp-files} Select top $k$ most suspicious classes, according to $R$;
  \item\label{step:compute-slice} Compute slicer for every suspicious
    file, obtained in Step~\ref{step:susp-files}, and every failing test;
  \item\label{step:group} Assemble all the resulting slicers of each failing test;
  \item Adjust spectra $S$, from step~\ref{step:spectra}, with
    the slices of each failing test, from
    step~\ref{step:group}, and then
	recompute the ranking.
\end{enumerate}

@@

%
\section{Evaluation}
\label{sec:eval}

We studied the following research questions:

\newcommand{\rqone}{How effective is DS in eliminating code?}
\newcommand{\rqtwo}{How often does DS miss faulty statements?}
\newcommand{\rqthree}{How effective is \comb{} for bug localization?}

\begin{itemize}[leftmargin=0em]
\item[]{\textbf{RQ1.}}~\textit{\rqone{}}
\item[]{\textbf{RQ2.}}~\textit{\rqtwo{}}
\item[]{\textbf{RQ2.}}~\textit{\rqthree{}}
\end{itemize}

The first question \Fix{...elaborate...}
The second question addresses the problem that affects all dynamic
slicing techniques~\cite{Lin:2018:BDE:3238147.3238163}---that of
missing faulty statements (see
Section~\ref{sec:slicing}). Understanding that issue is important; the
consequence for debugging is that developers would never be able to
locate faults. The third research question evaluates the impact of the
combination. To sum, our results show that the combination of the two
approaches improves the accuracy of fault localization on average by
\avgImprov\ in 51\% of the studied faults.

%%  shows that CS can
%% miss faulty statements because of imprecise oracles that guide the
%% slicing process. 
%% Different slicers can miss faulty statements for different
%% reasons. In this paper, we used Critical
%% Slicing~\cite{DeMillo:1996:CSS:229000.226310}, which, in principle,

%
To evaluate our approach, we compare the cost of diagnosing
a collection of faults using \sfl{} versus \comb{}.
%
\subsection{Objects of Analysis}\label{sec:analysis}
